---
title: "Discrete Time & Continious Time Event History Analysis"
author: "Sergej Kaiser"
date: "3/05/2017"
output:
  html_document: default
  html_notebook: default
css: markdown7.css
---
The document shows some of the visualizations, which I created for the course 'Event history analysis'. 
The focus of event history anlysis is to predict the time-to-event outcomes, such as e.g. birth of a child. <br>
The first visualizations show the discrete time analysis of the chance of a third birth. The continious time analysis is about general mortality risk.
```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
setwd("/Users/sergej/blog")
library(knitr)
library(tidyverse)
library(lazyeval)
library(forcats)
library(htmlTable)
library(texreg)
library(gridExtra)
library(grid)
if ("package:knitr" %in% search()){
  # Set some basic options. You usually do not
  # want your code, messages, warnings etc
  # to show in your actual manuscript
  opts_chunk$set(warning=FALSE, 
                 message=FALSE, 
                 echo=FALSE, 
                 dpi=96,
                 fig.width=12, fig.height=8, # Default figure widths
                 dev="png", dev.args=list(type="cairo"), # The png device
                 # Change to dev="postscript" if you want the EPS-files
                 # for submitting. Also remove the dev.args() as the postscript
                 # doesn't accept the type="cairo" argument.
                 error=FALSE)
  
 
}
```
```{r data}
library(readr)
third.birth <-readr::read_rds("event_hist_data_part1.rds")
mort.data <- readr::read_rds('event_data_part2.rds')
```

```{r functions}
my.ggplot.hazard<-function(df,x,y,z=NULL,major.breaks=c(seq(0.00,0.10,0.02)),minor.breaks=c(seq(0.01,0.1,0.02)),title.str="Plot of third birth Hazard",legend.pos="none",pal.str="Dark2",ylab="Hazard"){
  if(is.null(z))
    {
    ggplot(df, aes_string(x, y)) + geom_line() +  theme_bw()+
      #coord_cartesian(ylim=ylimits)+
    labs(y=ylab,x="Years since second birth",title=title.str)+
      scale_y_continuous(breaks=major.breaks,minor_breaks = minor.breaks)+ 
      scale_x_continuous(breaks=c(seq(0,20,2)),minor_breaks = c(seq(1,20,2))) + 
      theme_minimal() +
        theme(
              panel.grid.major.x = element_blank(),
              panel.grid.minor = element_blank(),
              legend.title = element_blank(),
              legend.justification = c(0, 1), 
              legend.position = "bottom",
              legend.background = element_blank(),
              legend.direction="horizontal",
              text = element_text(size=14,family = "Georgia"),
              plot.title = element_text(hjust=0.5,vjust=0.5,size = 20, margin = margin(b = 10)),
              plot.subtitle = element_text(size = 14, color = "darkslategrey", margin = margin(b = 25)),
              plot.caption = element_text(size = 11, margin = margin(t = 10), color = "grey70", hjust = 0))
    }
  else{
    ggplot(df, aes_string(x, y,color=z)) + 
      geom_line() +  theme_bw()+
      #coord_cartesian(ylim=ylimits)+
      scale_color_brewer(palette = pal.str)+
    labs(y=ylab,x="Years since second birth",title=title.str)+
      scale_y_continuous(breaks=major.breaks,minor_breaks = minor.breaks)+ 
      scale_x_continuous(breaks=c(seq(0,20,2)),minor_breaks = c(seq(1,20,2)))+
    theme_minimal() +
        theme(
              panel.grid.major.x = element_blank(),
              panel.grid.minor = element_blank(),
              legend.title = element_blank(),
              legend.justification = c(0, 1), 
              legend.position = "bottom",
              legend.background = element_blank(),
              legend.direction="horizontal",
              text = element_text(size=14,family = "Georgia"),
              plot.title = element_text(hjust=0.5,vjust=0.5,size = 20, margin = margin(b = 10)),
              plot.subtitle = element_text(size = 14, color = "darkslategrey", margin = margin(b = 25)),
              plot.caption = element_text(size = 11, margin = margin(t = 10), color = "grey70", hjust = 0)
              )
    }}
theme_geom_line<-theme_geom_line<-function(pal.str,title.str,major.breaks,minor.breaks,ylab,legend.pos,...){
  list(
  geom_line() , 
  theme_bw(),
  scale_color_brewer(palette = pal.str),
  labs(y=ylab,x="Years since second birth",title=title.str),
  scale_y_continuous(breaks=major.breaks,minor_breaks = minor.breaks),
  scale_x_continuous(breaks=c(seq(0,20,2)),minor_breaks = c(seq(1,20,2))),
   theme_minimal(),
        theme(
              panel.grid.major.x = element_blank(),
              panel.grid.minor = element_blank(),
              legend.title = element_blank(),
              legend.justification = c(0, 1), 
              legend.position = "bottom",
              legend.background = element_blank(),
              legend.direction="horizontal",
              text = element_text(size=14,family = "Georgia"),
              plot.title = element_text(hjust=0.5,vjust=0.5,size = 20, margin = margin(b = 10)),
              plot.subtitle = element_text(size = 14, color = "darkslategrey", margin = margin(b = 25)),
              plot.caption = element_text(size = 11, margin = margin(t = 10), color = "grey70", hjust = 0)))
  }
```
  
```{r 1a_lifetable, include=FALSE,cache=TRUE}
life.table<-purrr::map(seq(0,19,1),.f=function(i){
	 risk_set <- dplyr::filter(third.birth, EXPOSURE%in%i & EXPOSURE<=MAXEXP) %>% nrow(.)
	 events <- dplyr::filter(third.birth, EXPOSURE %in% i & BIRTH3%in%1) %>% nrow(.)
	 censored <- dplyr::filter(third.birth, EXPOSURE%in%i & BIRTH3%in%0 & EXPOSURE %in% MAXEXP) %>% nrow(.)
	 life.t<-cbind(risk_set,events,censored)
	 as.data.frame(life.t)
})
life.table<-data.table::rbindlist(life.table,use.names = T,idcol="Year")

life.table<-life.table %>% mutate(hazard=events/risk_set)
life.table <- life.table %>% mutate(survivor=cumprod(1-hazard))
life.table$estimated_risk_set[2:nrow(life.table)] <- c(life.table$risk_set[1] * cumprod(life.table$survivor))[1:19]
life.table$estimated_risk_set[1]<-life.table$risk_set[1]
  columnames<-c("Year","Risk Set","Events","Censored","Pr(T<sub>i</sub>=j &#8739; T<sub>i</sub> &#8805;
                j)","Pr(T<sub>i</sub>>j)","Est. Risk Set")
  col_names<-c("Year","Risk Set","Events","Censored","Pr(T_i=j | T_i >= j)","Pr(T_i>j)","Est. Risk Set)")
    tb<-htmlTable(txtRound(life.table, 3), 
              header =  columnames,
              # Add a column spanner
              rnames=FALSE,
              caption="Life table partiy progression to the third Child",
              align="cccccccc",
              col.rgroup = c("none", "#F7F7F7"),
              css.rgroup = "")
```

### Discrete Time models

The first visualization displays the (lack of) fit of various specifications of the baseline hazard of the chance of a third birth. <br> 
We model the chance of the third birth using a logistic regression. For each model the outcome is the birth of a third child and the dependent variable is time  (const linear, quadratic, etc.).   

```{r 2a hazard models ,echo=FALSE,cache=T}
third.birth$EXPOSURE_F <- factor(third.birth$EXPOSURE)

third.birth.sub<-dplyr::select(third.birth,EXPOSURE_F,EXPOSURE,NR,EXP_CAT4)
########## fit polynominals equation degree one till four
res.glm.fit.logits<-map(seq(1,4,1),.f=function(i){
  glm.fit <- glm(BIRTH3 ~ poly(EXPOSURE,degree=i,raw=T), data=third.birth, family=binomial(link="logit"))
  pred.logits.glm.fit <- predict.glm(glm.fit)
  pred.odds.glm.fit <- exp(pred.logits.glm.fit)
  pred.prob.glm.fit <- pred.odds.glm.fit/(1+pred.odds.glm.fit)
  glm.fit.plot <- ggplot(third.birth,aes_string("EXPOSURE","pred.prob.glm.fit"))+theme_geom_line(pal.str="Set1",title="Plot of third birth Hazard",major.breaks=c(seq(0,0.1,0.02)),minor.breaks=c(seq(0.01,0.1,0.02)),ylab="Hazard",legend.pos = "bottom")
  res.glm<-list(glm.fit,pred.logits.glm.fit,pred.prob.glm.fit,glm.fit.plot)
  names(res.glm)<-c(paste("fit",i,sep="_"),paste("logits",i,sep="_"),paste("prob",i,sep="_"),paste("glm.fit.plot",i,sep="_"))
  res.glm})


# Constant Specification
glm.const.ls<-list()
glm.const.ls$glm.const <- glm(BIRTH3 ~ 1, data=third.birth, family=binomial(link="logit"))
glm.const.ls$pred.logits.glm.const <- predict.glm(glm.const.ls$glm.const) # predicted logits
glm.const.ls$predo.glm.const <- exp(glm.const.ls$pred.logits.glm.const)  #predicted odds
glm.const.ls$pred.prob.glm.const <-glm.const.ls$predo.glm.const/(1+glm.const.ls$predo.glm.const)   #predicted probabilities
glm.const.ls$glm.constplot <- ggplot(third.birth, aes_string("EXPOSURE", "glm.const.ls$pred.prob.glm.const")) +
   geom_line() + 
  coord_cartesian(ylim=c(0,0.1))+
  labs(y="Hazard",x="Years since second birth",title="Plot of third birth Hazard")+
  scale_y_continuous(breaks=c(0.00,0.02,0.04,0.06,0.08,0.10),minor_breaks = seq(0.01,0.1,0.02))+ 
  scale_x_continuous(breaks=c(seq(0,20,2)),minor_breaks = c(seq(1,20,2))) + theme_minimal() +
        theme(
              panel.grid.major.x = element_blank(),
              panel.grid.minor = element_blank(),
              legend.title = element_blank(),
              legend.justification = c(0, 1), 
              legend.position = "bottom",
              legend.background = element_blank(),
              legend.direction="horizontal",
              text = element_text(size=14,family = "Georgia"),
              plot.title = element_text(hjust=0.5,vjust=0.5,size = 20, margin = margin(b = 10)),
              plot.subtitle = element_text(size = 14, color = "darkslategrey", margin = margin(b = 25)),
              plot.caption = element_text(size = 11, margin = margin(t = 10), color = "grey70", hjust = 0))

names(glm.const.ls)<-c(paste("fit","gen",sep="_"),paste("logits","gen",sep="_"),paste("prob","gen",sep="_"),paste("glm.fit.plot","gen",sep="_"))
#

############ General Specification#########
glm.gen.ls<-list()
glm.gen.ls$glm.gen <- glm(BIRTH3 ~ EXPOSURE_F, data=third.birth, family=binomial(link="logit"))
glm.gen.ls$pred.logits.glm.gen <- predict.glm(glm.gen.ls$glm.gen)
glm.gen.ls$pred.odds.glm.gen <- exp(glm.gen.ls$pred.logits.glm.gen)
glm.gen.ls$pred.pop.glm.gen <- glm.gen.ls$pred.odds.glm.gen/(1+glm.gen.ls$pred.odds.glm.gen)
glm.gen.ls$glm.gen.plot <- my.ggplot.hazard(third.birth,"EXPOSURE", "glm.gen.ls$pred.pop.glm.gen")

names(glm.gen.ls)<-c(paste("fit","gen",sep="_"),paste("logits","gen",sep="_"),paste("prob","gen",sep="_"),paste("glm.fit.plot","gen",sep="_"))


########### Step Function #############
third.birth$EXP_CAT4<-as.factor(third.birth$EXP_CAT4)
glm.step.ls<-list()
glm.step.ls$glm.step <- glm(BIRTH3~EXPOSURE+EXP_CAT4, data=third.birth, family=binomial(link = "logit"))
glm.step.ls$pred.logits.glm.step <- predict(glm.step.ls$glm.step)
glm.step.ls$predo.glm.step <- exp(glm.step.ls$pred.logits.glm.step)
glm.step.ls$pred.prob.glm.step <- glm.step.ls$predo.glm.step/(1+glm.step.ls$predo.glm.step)
glm.step.ls$plot<-ggplot(data=third.birth, aes_string(x="EXPOSURE",y="glm.step.ls$pred.prob.glm.step")) +
  geom_line() 
#glm.step.ls<-list(glm.step,pred.logits.glm.step,predo.glm.step,pred.prob.glm.step)
names(glm.step.ls)<-c(paste("fit","step",sep="_"),paste("logits","step",sep="_"),paste("prob","step",sep="_"),paste("glm.fit.plot","step",sep="_"))
res.glm.fit.logits$constant<-glm.const.ls
res.glm.fit.logits$step<-glm.step.ls
res.glm.fit.logits$general<-glm.gen.ls
names(res.glm.fit.logits)<-c("linear","quadratic","cubic","poly_4th_order","constant","step","general")

hazard.const.lin.quad <- ggplot(data=third.birth, aes(x=EXPOSURE)) +
  coord_cartesian(ylim=c(0,0.12))+
  scale_colour_manual(values=c('#984ea3','#377eb8','#a65628','#000000'))+
  geom_line(aes( y=res.glm.fit.logits[[5]][[3]], color="Constant")) +
        geom_line(aes(y=res.glm.fit.logits[[1]][[3]],color="Linear")) +
			 geom_line(aes( y=res.glm.fit.logits[[2]][[3]], color="Quadratic")) +
			 geom_point(data=life.table, aes(x=0:19, y=hazard, color="Sample hazard"),size=1.25)+
			   labs(y="Hazard",x="Years since second birth")+
  #guides(guide_colorbar(title="Various specifications"))+
  scale_y_continuous(breaks=c(seq(0,0.12,0.02)),minor_breaks = seq(0.01,0.12,0.02))+ 
  scale_x_continuous(breaks=c(seq(0,20,2)),minor_breaks = c(seq(1,20,2)))+
  guides(colour=guide_legend(title="Specification",nrow=2))+
  theme_minimal() +
        theme(
              panel.grid.major.x = element_blank(),
              panel.grid.minor = element_blank(),
              legend.title = element_blank(),
              legend.justification = c(0, 1), 
              legend.position = "bottom",
              legend.background = element_blank(),
              legend.direction="horizontal",
              text = element_text(size=14,family = "Georgia"),
              plot.title = element_text(hjust=0.5,vjust=0.5,size = 20, margin = margin(b = 10)),
              plot.subtitle = element_text(size = 14, color = "darkslategrey", margin = margin(b = 25)),
              plot.caption = element_text(size = 11, margin = margin(t = 10), color = "grey70", hjust = 0))

hazard.cubic.four.step<-ggplot(data=third.birth, aes(x=EXPOSURE)) +
  coord_cartesian(ylim=c(0,0.12))+
  scale_colour_manual(values=c('#e41a1c','#ff7f00',"#000000",'#4daf4a'))+
  
			 geom_line(aes( y=res.glm.fit.logits[[3]][[3]], color="Cubic"))+ 
			 geom_line(aes( y=res.glm.fit.logits[[4]][[3]], color="Quartic")) +
  geom_line(aes( y=res.glm.fit.logits[[6]][[3]], color="Stepwise"))+
			 geom_point(data=life.table, aes(x=0:19, y=hazard,color="Sample hazard"),size=1.25)+
			   labs(y="Hazard",x="Years since second birth")+
  #guides(guide_colorbar(title="Various specifications"))+
  scale_y_continuous(breaks=c(seq(0,0.12,0.02)),minor_breaks = seq(0.01,0.12,0.02))+ 
  scale_x_continuous(breaks=c(seq(0,20,2)),minor_breaks = c(seq(1,20,2)))+
    guides(colour=guide_legend(title="Specification",nrow=2,override.aes = list(linetype=0,shape=c(15,15,16,15),size=3)))+ theme_minimal() +
        theme(
              panel.grid.major.x = element_blank(),
              panel.grid.minor = element_blank(),
              legend.title = element_blank(),
              legend.justification = c(0, 1), 
              legend.position = "bottom",
              legend.background = element_blank(),
              legend.direction="horizontal",
              text = element_text(size=14,family = "Georgia"),
              plot.title = element_text(hjust=0.5,vjust=0.5,size = 20, margin = margin(b = 10)),
              plot.subtitle = element_text(size = 14, color = "darkslategrey", margin = margin(b = 25)),
              plot.caption = element_text(size = 11, margin = margin(t = 10), color = "grey70", hjust = 1))

```

```{r disc hazards}
grid.arrange(hazard.const.lin.quad,hazard.cubic.four.step, nrow=1,ncol=2,top="Various baseline specification")
```


## Continious Event History 

### Kaplan Meier cumulative hazard

The Kaplan-Meier estimator is a simple non-parametric method to obtain an estimate of the survival curve from time-to-event data. 
It can be used with both discrete and continious time-to-event data. Here our focus is to model the outcome general mortality by educational categories.

#### Kaplan-Meier with covariate education

```{r 2d kaplan meier educ, cache=T}
library(survival)
kpm.educ.data<-mort.data
#kpm.educ.data$EDUCATOIN<-as_factor(kpm.educ.data$EDUCATION)
kp.meier.educ <- survfit(Surv(MONTH_CONT, DEATH) ~ EDUCATION, data = mort.data)
kp.meier.educ.sum<-summary(kp.meier.educ)
#survival function
kpm.educ.plot.data<-data.frame(kp.meier.educ.sum$time,kp.meier.educ.sum$surv,kp.meier.educ.sum$strata,kp.meier.educ.sum$lower,kp.meier.educ.sum$upper)
names(kpm.educ.plot.data)<-gsub("kp.meier.educ.sum.","",names(kpm.educ.plot.data))
levels(kpm.educ.plot.data$strata)<-c("Unknown","None & Primary","Lower Secundary","Higher Secundary","Higher Education")
 
km.surv.plot<-ggplot(kpm.educ.plot.data,aes(x=time,y=surv,color=strata))+scale_x_continuous(breaks=c(seq(0,36,4)),minor_breaks=c(seq(0,36,2)))+geom_line()+ scale_color_brewer(palette = "Dark2")+
  labs(x="Exposure",y="Survival Prob", title="Survival function")+guides(color=guide_legend("Education",reverse = T))+  theme_minimal() +
     theme(
              panel.grid.major.x = element_blank(),
              panel.grid.minor = element_blank(),
              legend.title = element_blank(),
              legend.justification = c(0, 1), 
              legend.position = "bottom",
              legend.background = element_blank(),
              legend.direction="horizontal",
              text = element_text(size=14,family = "Georgia"),
              plot.title = element_text(hjust=0.5,vjust=0.5,size = 20, margin = margin(b = 10)),
              plot.subtitle = element_text(size = 14, color = "darkslategrey", margin = margin(b = 25)),
              plot.caption = element_text(size = 11, margin = margin(t = 10), color = "grey70", hjust = 1))
              
```


```{r cum hazard kpm, echo=FALSE, message=FALSE, warning=FALSE, cache=T}
#
#plot(kp.meier.educ, fun="cumhaz", col=c(1:10), main="Cumulative hazard function")
#legend(x = "topleft", legend=labels.pl, col=c(1:10), lty=c(1,1))


cum.haz.df <- kpm.educ.plot.data
cum.haz.df$l.cum.haz<-log(-log(cum.haz.df$surv))
cum.haz.df$cum.haz<--log(cum.haz.df$surv)
#levels(cum.haz.df$educ)<-gsub("EDUCATION=","",levels(cum.haz.df$educ))
p.cum.haz <- ggplot(cum.haz.df, aes(time,cum.haz,color=strata)) +scale_x_continuous(breaks=c(seq(0,36,4)),minor_breaks=c(seq(0,36,2)))+scale_y_continuous(breaks=c(seq(0.0,0.12,0.02)),minor_breaks = c(seq(0.01,0.12,0.02)))+ labs(x="Exposure",y="Cumulative Hazard",title="Cumulative Hazard")+geom_step() +scale_color_brewer(palette = "Dark2")+guides(color=guide_legend(title="Education"))        + theme_minimal() +
         theme(
              panel.grid.major.x = element_blank(),
              panel.grid.minor = element_blank(),
              legend.title = element_blank(),
              legend.justification = c(0, 1), 
              legend.position = "bottom",
              legend.background = element_blank(),
              legend.direction="horizontal",
              text = element_text(size=14,family = "Georgia"),
              plot.title = element_text(hjust=0.5,vjust=0.5,size = 20, margin = margin(b = 10)),
              plot.subtitle = element_text(size = 14, color = "darkslategrey", margin = margin(b = 25)),
              plot.caption = element_text(size = 11, margin = margin(t = 10), color = "grey70", hjust = 1))
              
#cox.age
grid_arrange_shared_legend <- function(..., ncol = length(list(...)), nrow = 1, position = c("bottom", "right")) {

  plots <- list(...)
  position <- match.arg(position)
  g <- ggplotGrob(plots[[1]] + theme(legend.position = position))$grobs
  legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
  lheight <- sum(legend$height)
  lwidth <- sum(legend$width)
  gl <- lapply(plots, function(x) x + theme(legend.position="none"))
  gl <- c(gl, ncol = ncol, nrow = nrow)

  combined <- switch(position,
                     "bottom" = arrangeGrob(do.call(arrangeGrob, gl),
                                            legend,
                                            ncol = 1,
                                            heights = unit.c(unit(1, "npc") - lheight, lheight)),
                     "right" = arrangeGrob(do.call(arrangeGrob, gl),
                                           legend,
                                           ncol = 2,
                                           widths = unit.c(unit(1, "npc") - lwidth, lwidth)))

  grid.newpage()
  grid.draw(combined)

  # return gtable invisibly
  invisible(combined)

}
```
```{r grid, message=F,echo=F}
grid_arrange_shared_legend(km.surv.plot,p.cum.haz, position ="bottom")
```

### Cox regression with covariate education

We continue to study the outcome general mortality risk. The following graph shows the recovered non-parametric cummulative hazard of general mortality by educational according. We obtained the cum. hazard function from a cox regression. <br>
The cox regression is a tool to build predictive models for continious time-to-event data (see: e.g. SPSS Manual 22). It allows to study the effect of a covariate on the outcome without assumptions about the hazard function. <br>
<!-- The figure shows the effect of education on the estimated survival function.  -->

```{r 2.e COX regression EDUCATOIN, echo=FALSE, message=FALSE, warning=FALSE, cache=T, results='asis'}
mort.data$EDUCATION<- haven::as_factor(mort.data$EDUCATION)

mort.data$HOUSING <- haven::as_factor(mort.data$HOUSING)

cox.educ.5<- coxph(Surv(mort.data$MONTH_CONT, mort.data$DEATH) ~ EDUCATION, data = mort.data)

#stargazer(cox.educ.5,type="html",dep.var.labels = "Log H(t)",covariate.labels=c(paste0(paste0("&beta;<sub>",seq(2,5,1)),"</sub>")),keep.stat=c("n","ll"))
wald.test.stats<-round((cox.educ.5$coefficients)^2/diag(cox.educ.5$var),3)
```

```{r cox educ survival plot,cache=T}

new_df <- with(mort.data,
               data.frame(EDUCATION = c(levels(haven::as_factor(mort.data$EDUCATION)))
                          )
)

# Survival curves with new data
#%%%%%%%%%%%%%%%%%%%%%%%%%%%
#Approach taken from https://github.com/kassambara/survminer/issues/67
library(survminer)
library(survival)
fit.cox.educ.5 <- survfit(cox.educ.5, newdata = new_df)
cox.educ.fit5<-survfit(cox.educ.5)
# plot<-ggsurvplot(fit.cox.educ.5, conf.int = F, palette = "Dark2",ylim=c(0.9,1), break.time.by = 4, ggtheme=theme_bw(),xlab = "Months",y="Survival",main = "Cox Regression within group survival")
# plot$plot+scale_color_brewer(palette="Dark2",labels =c(levels(mort.data$EDUCATION)))+guides(color=guide_legend(title="Education")) +theme(plot.title = element_text(hjust=0.5,vjust=0.5),legend.position = "bottom")
```

```{r 2.e COX Educ cumhaz,cache=T}
#autoplot(survfit(cox.educ.5),fun = 'cumhaz')
#obtain recoverd baseline cum.hazard
plot.cumhaz.eudc.cox<-ggsurvplot(fit.cox.educ.5,font.x=11,font.y=11,font.main=12, conf.int = F, linetype="solid",palette = "Dark2",ylim=c(0.0,0.12), break.time.by = 4, ggtheme=theme_minimal(),xlab = "Months",ylab="Cum Hazard",main = "Cox Regression within group cumulative hazard",fun="cumhaz",risk.table = F)
plot.cumhaz.eudc.cox$plot+scale_color_brewer(palette="Set1",labels =c(levels(mort.data$EDUCATION)))+guides(color=guide_legend(title="Education"))        +
        theme(
        
              panel.grid.major.x = element_blank(),
              panel.grid.minor = element_blank(),
              legend.title = element_blank(),
              legend.justification = c(0, 1), 
              legend.position = "bottom",
              legend.background = element_blank(),
              legend.direction="horizontal",
              text = element_text(size=14,family = "Georgia"),
              plot.title = element_text(hjust=0.5,vjust=0.5,size = 20, margin = margin(b = 10)),
              plot.subtitle = element_text(size = 14, color = "darkslategrey", margin = margin(b = 25)),
              plot.caption = element_text(size = 11, margin = margin(t = 10), color = "grey70", hjust = 0))
```  